Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea

### 2022-10-15 analysis/paper/paper.qmd
editing the abstract and intro

abstract: |
 The introduction of blade technology, stemmed points, end scrapers, burins, denticulates, and finer grained materials led to the transition from the Early to Late Paleolithic in Korea. Stemmed points have been considered a representative tool that led this whole set of changes. In this research, we examine the function of stemmed points to understand the role that they played during the technological transition as well as throughout the Late Paleolithic. Our main questions are: What were stemmed points used for? How diverse were their functions? What are the temporal patterns in stemmed point functions? We measured tip cross-sectional areas (TCSA) to discriminate different functional classes of projectile points, for example, poisoned arrowheads or thrusting spear. We analyze TCSA with other variables including raw materials, weight, site and radiocarbon dates. Our results show that the stemmed points mostly functioned as javelins and thrusting spear tips, with smaller numbers as dart tips and arrowheads. TCSA values are depending mostly on size rather than raw material types. We found different usage of stemmed points in different sites, which could indicate people used stemmed points in different ways depending on the environment. However, some sites show a wide range of TCSA values that represent multi-purpose usage of stemmed points. The temporal pattern of TCSA values is one of little change throughout the Late Paleolithic period. We conclude that stemmed points were mainly designed for Javelin and thrusting spear tips but they were multi-functional tools.
keywords: |
  Stemmed points; Korean Paleolithic; Tip Cross-Sectional Area
---

@@

With the introduction of new technologies, the Korean Paleolithic transitioned from the Early to the Late Paleolithic periods. The transition includes blade technology, stemmed points, end scrapers, burins, denticulates, etc [@bae2017origin; @nakazawa2018quaternary; @bae2017late; @lee2017korean; @seong2016eastern]. Another notable change is the selective use of raw materials along with the emergence of new tools. Previously quartzite and vein quartz were the most commonly used for core and flake tools but finer grained materials such as silicified tuff (shale), chert, hornfels, and obsidian became more important to the lithic technology during the Late Paleolithic. While people still used coarse materials with existing tools, they selectively chose finer materials for newly introduced tools. [@Seong_2003; @seong2004quartzite]. Stemmed points are considered the first evidence of the new technology, as well as the Late Paleolithic, since they led to the whole set of changes [@seong2016eastern; @seong2008tanged]. This is related to the fact that stemmed points originate from Korea in Northeast Asia and have a close association with mobility, site formation, and occupation diversity [@o2018origins; @Chong_2021; @park2022did]. Despite the importance of stemmed points, only a few studies have examined their use, whereas researchers often discuss stemmed points relating to their origin and chronology of the Korean Late Paleolithic and their relationship with Japan [@chang2013human; @Park_2013; @lee2019were; @Chong_2021].[@chang2013human; @Park_2013; @lee2019were; @Chong_2021].

>>>>

abstract: |
 The introduction of blade technology, stemmed points, end scrapers, burins, denticulates, and finer grained materials characterises the transition from the Early to Late Paleolithic in Korea. Stemmed points have been considered a representative tool that led this set of changes. In this research, we examine the possible role that stemmed points played during the technological transition, as well as throughout the Late Paleolithic. Our main questions are: What were stemmed points used for? How diverse were their likely uses? What are the temporal patterns in stemmed point use? We measured tip cross-sectional area (TCSA) to discriminate different likely use classes of projectile points, for example, as poisoned arrowheads or as thrusting spears. We analyzed TCSA with other variables including raw materials, weight, site and radiocarbon dates. Our results show that the stemmed points were likely used  as javelins and thrusting spear tips, with smaller numbers as dart tips and arrowheads. TCSA values are controlled mostly by size rather than raw material types. We found different usage of stemmed points at different sites, which could indicate people used stemmed points in different ways depending on the local environment. Some sites show a wide range of TCSA values that represent multi-purpose usage of stemmed points. The temporal pattern of TCSA values is one of little change throughout the Late Paleolithic period. We conclude that stemmed points were  multi-functional tools, with many likely designed for javelin and thrusting spear tips.
keywords: |
  Stemmed points; Korean Paleolithic; Tip Cross-Sectional Area
---

@@

The introduction of new stone artefact technologies marks a major transition in the Korean Paleolithic from the Early to the Late Paleolithic periods. The transition includes the appearance blade technology, stemmed points, end scrapers, burins, denticulates, etc [@bae2017origin; @nakazawa2018quaternary; @bae2017late; @lee2017korean; @seong2016eastern]. Stemmed points are considered the first evidence of a suite of new technologies defining the Late Paleolithic in this region [@seong2016eastern; @seong2008tanged]. This is related to the fact that stemmed points originate from Korea and spread throughout Northeast Asia, and they have a close association with mobility, site formation, and occupation diversity [@o2018origins; @Chong_2021; @park2022did]. Despite the importance of stemmed points, only a few studies to date have examined their likely uses. Previous work has mostly debated their origin, the chronology of the Korean Late Paleolithic, and their relationship with Japan [@chang2013human; @Park_2013; @lee2019were; @Chong_2021].[@chang2013human; @Park_2013; @lee2019were; @Chong_2021].
